The framing of active school travel in Ontario, Canada as a health and built environment issue

### 2021-11-26 AST-Framing-Ontario/AST-Framing-Ontario.Rmd

---
title: How do school travel planning stakeholders frame active school travel in Ontario, Canada?
author:
  - name: Author 1
    email: author1@example.com
    affiliation: Some Department
    footnote: 1
  - name: Author 2
    email: author2@example.com
    affiliation: Some Department
  - name: Author 3
    email: author3@example.com
    affiliation: Another University
    footnote: 2
  - name: Author 4
    email: author4@example.com
    affiliation: Some Institute
    footnote: 2
  - name: Author 5
    email: author5@example.com
    affiliation: Some University
  - name: Author 6
    email: author6@example.com
    affiliation: Some Department
address:
  - code: Some Department
    address: Department, Street, City, Province, Postal Code
  - code: Another University
    address: Department, Street, City, Province, Postal Code
  - code: Some Institute
    address: Street, City, Province, Postal Code
  - code: Some University
    address: Department, Street, City, Province, Postal Code
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
abstract: |
  This is the abstract.

@@

journal: "Journal of Transport & Health"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
linenumbers: false
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

@@

Walking and bicycling to school, commonly known as active school travel (AST), has been declining in Canada and North America for decades [@rothmanDeclineActiveSchool2018], with levels much lower than other developed countries like The Netherlands [@vangoeverdenSchoolTravelBehaviour2013] and Japan [@waygoodChapterSixteenJapan2020]. In response to this trend, school travel planning has become a popular intervention across Canada that has resulted in mixed success with some schools reporting an increase in AST [e.g., @buliungSchoolTravelPlanning2011; @mammenActiveSchoolTravel2014], while others did not observe a change in behaviour [@buttazzoniPromotingActiveSchool2019] or needed more time and resources to be effective [@mammenPuttingSchoolTravel2015].

School travel planning (STP) is a "school-specific" intervention led by a facilitator that brings together a committee of stakeholders from diverse sectors including education, planning, transportation, and public health to develop action plans [@mammenSchoolTravelPlanning2014]. This five-step process involves identifying barriers to AST based on the local context and implementing approaches or activities that alleviate concerns and encourage AST [@faulknerSchoolTravelPlans2018]. STP facilitators and committee members generally intervene at the school through a range of activities such as educational strategies, in-person events or programs, improvements to the school site, and enforcement [@mammenActiveSchoolTravel2014].

How STP is framed to a target audience may influence how this intervention is received and whether it encourages the audience to support it or participate. In particular, STP stakeholders make choices about the points of intervention and potential benefits that ought to be communicated to the public to facilitate adoption of AST. Educational strategies, which account for approximately one quarter of STP strategies implemented at Canadian schools [@mammenActiveSchoolTravel2014], need to effectively engage multiple audiences – parents, students, politicians – on the issue, challenges, and benefits of AST. Promotional materials, therefore, should reflect current knowledge from research about school travel and the built environment, as well as local content specific to factors that influence travel decisions for parents, in order to effectively engage the public in advocating for improving AST or to expose parents and the community to essential information that may prompt behaviour change.

On the one hand, parents have been found to express different understandings, nomenclature, and conceptualizations than planners of the factors that influence school travel [@buliungLivingJourneySchool2021]. However, it is unknown to what degree there is concurrence between STP materials presented to the public and the literature on school travel and parental perceptions, and how the complexity of the AST issue is presented to the public.

The aim of this paper is to fill this research gap by analyzing how AST is framed by STP stakeholders in Ontario, Canada. We used text mining and topic modelling to examine how local policy documents from Ontario municipalities and school boards present and communicate the benefits and barriers of AST and the solutions for improving AST. We compared the findings from these documents to a selection of studies on AST and explored the extent to which there is concurrence between research findings to date and promotional materials shared with the public to increase AST.

@@

@@

We assembled a collection of publicly available documents that were sourced online from the main stakeholder groups involved in STP initiatives in Ontario: i) school boards (public and English-speaking only); ii) municipal governments; and iii) transportation consortia. Non-profit organizations, police services, and advocacy groups are other stakeholders who may play a role in supporting AST and/or STP, but this study does not include any documents from these groups because they are not consistently participating in all initiatives.

The search was guided first by a list of all English public school boards across Ontario. The websites of each school board were manually searched for pages related to school transportation or travel. Any pages relevant to these topics were manually downloaded. Next, we collected documents by searching municipal government and transportation consortia websites. The latter were identified based on geographic area (i.e., the municipalities and/or transportation consortia that are in the same geographic area of each school board). Pages related to active or school travel were manually downloaded. Webpages from STP stakeholder websites were included in our analysis if they were easy to find. This primary criteria was important since our analysis pertains to how such issues are framed to the public. Thus, we included only webpages that were easy to find, which we defined as requiring no more than 2-4 separate links from the initial Google search.

The initial corpus of policy documents included 64 relevant webpages (i.e., one page or more) from all STP stakeholder groups. It is important to note that school boards, municipalities, and transportation consortia may or may not publish information about their involvement in AST and STP efforts on their respective websites or in policy documents, which means that some of these groups for particular regions in Ontario are not included in our analysis. Search results are summarized in Table \ref{tab:policy-documents}.

```{r policy-documents, echo=FALSE}
data_tbl <- data.frame(
  Stakeholder = c("School boards", "Municipalities", "Transportation consortia"),
  Total = c("62", "62", "39"),
  Included = c("31", "25", "8")
)

@@

### 3.2.1. Academic papers

@@

A multi-step process was conducted to ensure that the analysis captured as much content as possible from both the policy documents (n = 64) and academic papers (n = 233). To begin, the webpages, which were manually downloaded in portable document format (PDF), were trimmed so that pages that only consisted of tables, figures, or references were removed. Many academic papers were in a two-column format, which is not ideal for conversion to `txt`. We adapted a procedure (https://stackoverflow.com/questions/42541849/extract-text-from-two-column-pdf-with-r) to read the two-column PDF documents so that they would be converted correctly. Four academic papers did not join sufficiently and were taken out of the corpus due to the substantial time required to manually correct their inconsistencies.

Next, we converted the trimmed PDF documents into `txt` files so that they could be imported in R for analysis. We then proceeded to a manual cleaning phase where we removed any remaining tables, figures, references, headers/footings, and captions that could not be trimmed. Manual corrections were also required for certain pages in academic papers that remained in two-column format after the conversion process. This typically occurred on pages that had a table or figure which disrupted the text. Finally, we reviewed all of the documents to remove hyphenation by line breaks and to keep hyphenated words together on the same line. Any ligatures (e.g., combinations of characters or letters that were not properly detected during the conversion process) were fixed by inserting the unicode sequence of character to replace the missing sequence of characters.

@@

Issues that pertain to public health or wellbeing are often presented to the public through particular frames to influence perceptions or behaviours. Gamson and Modigliani [-@gamsonChangingCulture1987] conceptualize a frame as a “central organizing idea or story line that provides meaning” to a particular phenomenon. A frame can enable individuals "to locate, perceive, identify, and label" information pertaining to various dimensions of an issue [@goffmanFrameAnalysisEssay1974]. Scholars in the field of political communications have proposed that communicators, such as the media or an institution, construct the narrative of a frame for policy positions or public issues in order to activate or restrict a particular response in the intended audience [@panFramingAnalysisApproach1993]. Framing can be used to position existing solutions as suitable to address particular issues [@mahFramecriticalPolicyAnalysis2014], which may prevent the public from being aware of other policy approaches that challenge the status quo to bring about more radical change. The way policy issues are framed is ultimately important to understand because it plays a role in either altering or preserving the existing social perceptions.

Framing of issues is an important step in developing health policy. An obvious example over the past decade is the framing of climate change as a public health issue [e.g., @ depouxCommunicatingClimateChange2017; @maibachReframingClimateChange2010; @weathersDevelopmentsFramingClimate2016] to increase public engagement and reception of the issue. Transport planners also use different frames to guide the extent to which transport policies can be adapted to address climate change. In a recent paper [@reynardGrowthResilienceHow2021], framing analysis was applied to review the representation of issues such as mobility and social exclusion in municipal policies from four western Canadian cities under the current circumstances of climate change. The authors found four primary frames: “The Growing City”, “If You Build It, They Will Come”, “Better City for All”, and a “the Resilient City”. Each frame presented the nature, opportunities, and challenges of climate change in different ways which set the stage for the types of mitigation and adaptation strategies that cities were proposing to address this issue.

Given the range of factors at multiple levels of the socioecological model that influence school travel, it is important to understand how walking and bicycling to school are framed to the public. Is active school travel framed as a public health or environmental issue? Is it framed as both? Do proposed solutions to increase AST focus primarily on the intrapersonal and interpersonal factors, meaning that behaviour change must ultimately result from different decisions made by the individual or the household? Or is the decline in AST framed as an issue that must be addressed through changes in the built environment? These are the types of questions that we aim to answer in this paper. To our knowledge, this topic has not be addressed in the school travel literature to date.

@@

The topic modelling analysis was conducted in R. We primarily used the following packages: `topicmodels` [@R-topicmodels], `word2vec` [@R-word2vec], and `wordcloud` [@R-wordcloud] to examine content in the documents that were sourced for this project. These packages have functions for determining the frequency of specific words in each document, relationships (e.g., pairs of adjacent terms) and correlations between words, and estimating Latent Dirichlet allocation (LDA) models to classify documents according to the topics that are contained within them. We describe in the section below how we used these functions to produce our results.

@@

We analyzed word and document frequency for each corpora. Table \ref{tab:word-table} shows the most frequent terms found in the municipal, transportation consortia, school board, and academic documents. As expected, STP documents and academic papers reference _active travel_, _walking_, _biking_ or _cycling_, and _students_ more than other terms. Each corpora also has _safety_ and _traffic_ as common words which suggests congruence on these key factors between the research literature and how AST is framed to the public. The word _physical_ is present in each corpus which could refer to either _physical activity_, _physical health_, or the _physical environment_. Furthermore, documents from STP stakeholders discuss _resources_, _information_, and _services_ about school travel. In the section below, the context in which these terms appear is explored further through their concordance. Unlike the academic papers, STP stakeholder documents include the words _route_ or _routes_. This could reflect their role in identifying safe routes to school to share with parents or families, as well as the STP emphasis on making the physical environment safer for AST. The academic corpora differs from the policy documents in that _parents_ and _distance_ are the second and third most common terms. In addition, _time_, _factors_, _environment_, and _age_ are also identified in academic papers.  These words are absent from the list of common words in policy documents. Table \ref{tab:word-table} indicates that the research corpora discusses a broader range of determinants of AST than the policy documents. The number of references for each term in the academic papers is also significantly higher due to the inclusion of more documents.

@@

Bigrams for each policy corpora that occur more than 5 times are shown in Figures \ref{fig:city-visual},  \ref{fig:consortia-visual}, and \ref{fig:school-visual}. These figures help to make further sense of the word frequencies reported above, and highlight the main ideas that are presented to the public in each of the policy corpora. Municipalities primarily discuss _physical activity_ (n = 53) and _public health_ (n = 19) in the context of active travel. In addition, _travel planning_ (n = 19), _bike lanes_ (n = 16), and _safe routes_ (n = 14) are also identified, conceivably as interventions and built environment factors that support AST. Key issues related to transport such as _traffic safety_ (n = 10), _air quality_ (n = 9), and _greenhouse gases_ (n = 9) are conveyed to the public through AST documents. It is not surprising to find this focus in AST documents given that municipalities in Ontario are concerned about climate change and have increasingly looked to active modes of travel to offset transport-related emissions in urban areas.

Similar word bigrams are found in school board documents: _travel planning_ (n = 33), _safe routes_ (n = 15), _physical activity_ (n = 10), and _public health_ (n = 10) are among the most common bigrams. Both municipalities and school boards in Ontario seem to emphasize what can be or has been done to improve AST (i.e., policy or planning changes), while outlining some of the benefits of AST at the individual- or community-level to potentially encourage behaviour change (i.e., physical activity for children or improved air quality). Unlike other STP stakeholders, school boards also consider _inclement weather_ (n = 24) and _bus cancellations_ (n = 13). This is likely because many students in Ontario travel to school by bus and this information is presented alongside AST options. Finally, transportation consortia documents highlight topics such as _physical activity_ (n = 10), _pedestrian safety_ (n = 8), _crossing guards_ (n = 6), _travel planning_ (n = 6), and _walk zones_ (n = 6). Biking or cycling is notably absent from transportation consortia documents.

@@

policy_lda <- LDA(policy_dtm, k = 5, method = "Gibbs", control = NULL, model = NULL)

>>>>

---
title: The framing of active school travel in Ontario, Canada as a health and built environment issue
author:
  - name: Author 1
    email: author1@example.com
    affiliation: Some Department
    footnote: 1
  - name: Author 2
    email: author2@example.com
    affiliation: Some Department
  - name: Author 3
    email: author3@example.com
    affiliation: Another University
    footnote: 2
  - name: Author 4
    email: author4@example.com
    affiliation: Some Institute
    footnote: 2
  - name: Author 5
    email: author5@example.com
    affiliation: Some University
  - name: Author 6
    email: author6@example.com
    affiliation: Some Department
address:
  - code: Some Department
    address: Department, Street, City, Province, Postal Code
  - code: Another University
    address: Department, Street, City, Province, Postal Code
  - code: Some Institute
    address: Street, City, Province, Postal Code
  - code: Some University
    address: Department, Street, City, Province, Postal Code
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
abstract: |
  This is the abstract.

@@

journal: "Social Science & Medicine"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
linenumbers: false
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

@@

Rates of walking and bicycling to school, commonly known as active school travel (AST), have been declining in Canada and the United States for decades [@rothmanDeclineActiveSchool2018], with levels much lower than other developed countries like The Netherlands [@vangoeverdenSchoolTravelBehaviour2013], Denmark [@jensenHowObtainHealthy2008], and Japan [@waygoodChapterSixteenJapan2020]. However, the appetite for AST may be high in many Canadian communities. For instance, 40% of children in a study conducted in Toronto, Ontario who are driven to school would like to travel by bicycle instead [@laroucheRatherBikeSchool2016a]. But the study reported that less than 3% actually do, despite the vast majority having access to a bicycle and living within a short and bikeable distance to school. Another study based in London, Ontario reported similar findings with respect to children's preference for active travel [@larsenRouteBasedAnalysisCapture2012]. From a public health perspective, this represents a huge opportunity to improve health and wellbeing if more children could use active modes to school.

In response to this trend, the Government of Ontario, in Canada, created a fund in 2017 to support communities across the province in developing AST initiatives. Green Communities Canada runs the program, called _Ontario Active School Travel_, and determines the projects that will be funded with their support [@GreenCommunities2016]. As of December 2021, OAST has funded and provided resources for over 20 projects across Ontario.  Many communities implemented school travel planning (STP), along with encouragement activities like walking school buses or developed resources for schools and parents, among other actions [see @GreenCommunitiesProjects]. In this context, STP is a popular "school-specific" intervention led by a facilitator that brings together a committee of stakeholders from diverse sectors including education, planning, transportation, and public health to develop action plans [@buliungSchoolTravelPlanning2011; @mammenSchoolTravelPlanning2014]. The five-step process involves identifying barriers to AST based on the local context and implementing approaches or activities following the 5 E's that alleviate concerns about AST and make AST safer and more convenient [@buliungSchoolTravelPlanning2011; @langUnderstandingModalChoice2011].

How AST is framed by STP stakeholders to a target audience, like parents or the general public, may raise awareness about the issue and influence how walking and bicycling to school are perceived. Gamson and Modigliani [-@gamsonChangingCulture1987] define a frame as a “central organizing idea or story line that provides meaning” to a particular phenomenon. A frame can enable individuals "to locate, perceive, identify, and label" information pertaining to various dimensions of an issue [@goffmanFrameAnalysisEssay1974]. Therefore, stakeholder groups involved in STP efforts can play a role in shaping public perception about AST in such a way that it attracts greater attention and becomes a "problem" that needs to be addressed through behaviour change or by new policies. We also contend that the ways in which the AST issue is framed can affect the policy process, as outlined in Kingdon's [-@kingdonAgendasAlternativesPublic1984] Multiple Streams Framework (MSF), through all three streams (i.e., problem, policy, and politics). This framing can ultimately impact financial contributions and potential solutions to address the decline of AST in Canada.

After significant investment of human and financial resources to boost rates of AST in Ontario over the past few years, we ask the following questions: How is AST framed? What benefits of AST are communicated to the public? What solutions are proposed to increase rates of AST? Do proposed solutions focus primarily on the intrapersonal and interpersonal factors, meaning that behaviour change must ultimately result from the individual or household making different travel decisions? Or is AST framed as an issue that must be addressed through changes in the built environment or at the policy level?

In this paper, we use text mining and topic modelling to examine how three STP stakeholder groups in Ontario (i.e., municipalities, school boards, and transportation consortia) frame the issue of AST. We assembled a corpus of texts from webpages that can be considered an important source of information for the general public or parents who are interested in AST. We examine word frequency, bigrams, and concordances in these selected documents, and also identify key topics presented by each stakeholder group. We then compare the findings from these documents to a selection of studies on AST and explore the extent to which there is concordance between the literature on AST and materials shared with the public.

@@

## 2.1. Benefits of active school travel

The benefits of AST are valuable information to communicate to the public in order to convey the importance of this issue. The desire to increase AST in Canada is certainly warranted - there is compelling evidence that children who actively commute to school earn physical and mental health benefits. Faulkner et al. [-@faulknerActiveSchoolTransport2009] concluded from their systematic review that children who travel on foot or by bicycle to school generally have higher levels of physical activity than their peers who are driven to school. However, Schoeppe et al. [-@schoeppeAssociationsChildrenActive2015] reported no association between AST and physical activity. This relationship could be dose-dependent, meaning that children would have to travel longer distances to accumulate physical activity. A walking distance of 1000-1600 metres to school has been found to contribute to overall levels of physical activity for boys [@faulknerSchoolTravelChildren2013]. The daily routine of travelling to school can be a good opportunity for children to regularly build physical activity into their schedule [@mitraIndependentMobilityMode2013]. Research has also shown that using active modes to school contributes to improvements in cardiovascular fitness [@borrestadExperiencesRandomisedControlled2012].

More recently, there has been literature exploring the link between transport and children's wellbeing [@waygoodTransportChildrenWellbeing2020], with relevant applications to the study of travel satisfaction [@vandenbergFactorsAffectingParental2020; @westmanChildrenTravelSchool2017]. Being driven to school reduces community interactions for children which may impact their social wellbeing [@waygoodChildrenTravelIncidental2015], whereas walking and bicycling give children opportunities to socialize with friends or siblings [@michailChildrenExperiencesTheir2021]. This is something that children highly value about travel [@zwertsHowChildrenView2010a] and indicates that social connections through travel are important for their wellbeing. AST also provides opportunities for children to engage with the natural environment [@fuscoUnderstandingChildrenPerceptions2012; @romeroChildrenExperiencesEnjoyment2015] which can improve their mental or emotional wellbeing.

## 2.2. Factors that influence active school travel and mode choice

Factors that influence AST have been presented and organized using a socio-ecological model (SEM) [@mitraIndependentMobilityMode2013] or systems model [@badlandDevelopmentSystemsModel2016] whereby children's travel behaviour is understood within the context of their household, social, neighbourhood, and policy environments. Individual characteristics of the child are also determinants. The SEM comes from the field of public health and is a useful framework for understanding complex health behaviours, including travel habits such as walking or bicycling to school, because it identifies multiple determinants that need to be addressed by interventions to facilitate behaviour change. The consensus in the literature is that various factors are interrelated and that interventions should target multiple levels in order to increase levels of healthy and physically active modes of travel to school [@mitraIndependentMobilityMode2013]. Rather than describing an extensive list of factors that influence AST and mode choice in Canada, which has been covered elsewhere [e.g., @mammenUnderstandingDriveEscort2012; @mitraIndependentMobilityMode2013; @rothmanDeclineActiveSchool2018; @wilsonUnderstandingChildParent2018], we briefly discuss a few potentially modifiable determinants that may be targeted for change through STP.

At the individual level, older child age is often associated with AST [@mammenUnderstandingDriveEscort2012; @starkExploringIndependentActive2018; @wilsonUnderstandingChildParent2018] as parents typically escort younger children. There is some evidence that gender is a determinant of AST [@larsenInfluencePhysicalEnvironment2009], with boys being more likely to travel using active modes than girls, although this is not a strong or consistent finding [@rothmanDeclineActiveSchool2018; @schoeppeAssociationsChildrenActive2015]. Children's mode choice to school is strongly influenced by their parents' travel behaviours and the complexity of their household's travel needs [@buliungLivingJourneySchool2021], which indicates that shifting parental perceptions and habits is important. Convenience and inclement weather have been cited by parents as barriers to AST [@buliungSchoolTravelPlanning2011]. Parental perceptions of the built or school environment [@demeesterParentalPerceivedNeighborhood2014; @panterAttitudesSocialSupport2010] and their children's skills [@ghekiereInsightsChildrenIndependent2017] also influence whether they allow their children to walk or bicycle to school.

Distance between home and school is most strongly associated with AST [@ikedaAssociationsChildrenActive2018; @mammenUnderstandingDriveEscort2012; @pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018] with less AST reported among children who have to travel farther to school. Many studies have also found that the quality of the built environment along the route to school and around the school site [@ikedaAssociationsChildrenActive2018; @kimBuiltNaturalEnvironmental2020; @rothmanActiveSchoolTransportation2021] and provision of active travel infrastructure [@chenPromotingActiveStudent2018a; @pontEnvironmentalCorrelatesChildren2009] facilitate AST. Canadian youth report that they feel most safe bicycling on streets in their neighbourhood or that have low volumes of traffic [@TACbikeinfra2020]. Finally, concerns about traffic and strangers have been reported by parents who drive their children to school [@mammenUnderstandingDriveEscort2012], which highlights that the volume and speed of cars can be a concern or deterrent for AST.

## 2.3. School travel planning in Canada

School travel planning (STP) has been implemented in Canada since at least the late 2000s. Within the STP process, facilitators generally establish multi-sector committees who intervene at the participating school through a range of activities related to the 5E's such as _education_ strategies, _encouragement_ through in-person events or programs, _engineering_ improvements to or around the school site, and _enforcement_ of traffic speeds around schools [@langUnderstandingModalChoice2011; @mammenActiveSchoolTravel2014]. The first large-scale evaluation of STP as an intervention in Canada took place at twelve schools across the country, including four in Ontario, using parental surveys to measure changes in travel behaviour and perceptions [@buliungSchoolTravelPlanning2011]. Since then, there have been other assessments of both the efficacy of STP [@buttazzoniPromotingActiveSchool2019; @mammenActiveSchoolTravel2014] and the process of implementing such programs in Ontario [@buttazzoniSupportingActiveSchool2018; @mammenPuttingSchoolTravel2015]. STP facilitators have recommended that additional time and resources are needed to improve the efficacy of STP [@mammenPuttingSchoolTravel2015], which highlights that long-term and sustained efforts driven at the policy level are required to address declining rates of AST.

The ways in which AST is framed seem particularly important to shift parental attitudes and perceptions given their reported influence on children's travel mode to school. STP activities heavily focus on education or encouragement [@buliungSchoolTravelPlanning2011; @mammenActiveSchoolTravel2014; @buttazzoniSupportingActiveSchool2018], but parents may not always be receptive to the goals of STP and may be resistant to behaviour change [see @buttazzoniSupportingActiveSchool2018]. Parents have been found to express different understandings, language, and perceptions than planners of the built environment determinants that influence school travel [@buliungLivingJourneySchool2021]. This is also true when it comes to other factors like convenience of different modes to school [@langUnderstandingModalChoice2011]. STP stakeholder groups must pay special attention to parents' understanding of the decline of AST as a problem, which may affect their receptivity to proposed solutions.

The “central organizing idea or story line" of AST, to apply the definition from Gamson and Modigliani [-@gamsonChangingCulture1987], could also affect broader support in the community. Municipal representatives are perceived to be instrumental but the involvement of other stakeholder groups (e.g., busing consortium representatives and local residents) can be lacking [@buttazzoniSupportingActiveSchool2018]. In Ontario, it would be reasonable to say that AST has become a policy issue on the education and public health agendas over this time as evidenced by the financial contribution from the provincial government. The support from a range of municipal representatives [see @buttazzoniSupportingActiveSchool2018; @mammenPuttingSchoolTravel2015] demonstrates that the "policy stream" [see @kingdonAgendasAlternativesPublic1984] has been well engaged. It is unknown to what degree the general public (i.e., local residents) has been exposed to messaging about AST, which could affect their participation in the STP process as desired by STP facilitators [@buttazzoniSupportingActiveSchool2018].

The success of STP interventions would likely depend on parental judgments of factors that are related to school travel, as well as support from key policy makers. For example, parents have been found to view mixed land use as conducive for driving, despite transport planners viewing neighbourhoods with mixed uses as key for encouraging more active travel [@buliungLivingJourneySchool2021]. Therefore, stakeholder groups involved in STP must make important choices about the proposed solutions and potential benefits that ought to be communicated to parents and the general public about AST to convey its importance as a policy issue and to facilitate adoption of AST. Publicly available content about AST, therefore, needs to effectively engage multiple audiences on this policy issue including parents, children, politicians, and school representatives. This information should reflect current knowledge from research on school travel, plus content specific to local factors that influence AST, so that the challenges of AST are adequately defined and the opportunities or solutions to address the problems are clear. However, it is unknown to what degree STP materials presented to the general public reflect the scope of evidence from the literature on AST.

@@

We assembled a collection of publicly available documents that were sourced online from the main stakeholder groups involved in STP initiatives in Ontario: i) school boards (public or Catholic and English-speaking only); ii) municipal governments; and iii) transportation consortia. The latter are a unique group of entities sanctioned by Ontario's Ministry of Education in Ontario 2006. Each consortium involves a collaboration between regions and school boards; the consortium's objective is to deliver more efficient and timely transportation services to schools in each region [@ministryeducation2006]. Non-profit organizations, police services, and advocacy groups are other stakeholders who often play a role in supporting AST and/or STP, but this study does not include any documents from these groups because they are not consistently participating in all initiatives across Ontario.

The search was guided first by a list of all English public and Catholic school boards across Ontario. The websites of each school board were manually searched for pages related to school transport or travel. Any pages relevant to these topics were manually downloaded. Next, we collected documents by searching municipal government and transportation consortia websites. These were identified based on geographic area (i.e., the municipalities and/or transportation consortia that are in the same geographic area of each school board). Webpages related to active transport or school travel were manually downloaded.

Webpages from STP stakeholder groups were included in our analysis if they were easy to find. This primary criterion was important since our analysis pertains to how such issues are framed to the general public. Thus, we included only webpages that were readily accessible, which we defined as requiring no more than 2-4 separate links from the initial Google search.

The initial corpus of documents from STP stakeholder groups included 69 relevant webpages (i.e., one page or more) from all STP stakeholder groups. We refer to these as policy documents throughout the paper. It is important to note that school boards, municipalities, and transportation consortia may or may not publish information about their involvement in AST and STP efforts on their respective websites or in policy documents. Search results are summarized in Table \ref{tab:policy-documents}.

```{r policy-documents, echo=FALSE}
data_tbl <- data.frame(
  Stakeholder = c("School boards", "Municipalities", "Transportation consortia"),
  Total = c("62", "62", "39"),
  Included = c("32", "28", "9")
)

@@

### 3.1.2. Academic papers

@@

A multi-step process was conducted to ensure that the analysis captured as much text as possible from both the policy documents (n = 64) and academic papers (n = 233). To begin, the webpages, which were manually downloaded in portable document format (PDF), were trimmed so that pages that only consisted of tables, figures, or references were removed. Many academic papers were in a two-column format, which is not ideal for conversion to `txt`. We adapted a procedure (https://stackoverflow.com/questions/42541849/extract-text-from-two-column-pdf-with-r) to read the two-column PDF documents so that they would be converted correctly. Four academic papers did not join sufficiently and were taken out of the corpus due to the substantial time required to manually correct their inconsistencies.

Next, we converted the trimmed PDF documents into `txt` files so that they could be imported in R for analysis. We then proceeded to a manual cleaning phase where we removed any remaining tables, figures, references, headers/footings, and captions that could not be trimmed. Manual corrections were also required for certain pages in academic papers that remained in two-column format after the conversion process. This typically occurred on pages that had a table or figure which disrupted the text. Finally, we reviewed all of the documents to remove hyphenation by line breaks and to keep hyphenated words together on the same line. Any ligatures (e.g., combinations of characters or letters that were not properly detected during the conversion process) were fixed by replacing the unicode sequence of character by inserting the missing sequence of characters.

@@

Issues that pertain to public health or wellbeing are often presented to the public through particular frames to influence perceptions or behaviours. As previously mentioned, a frame is a "central organizing idea or story line that provides meaning" to a public issue or phenomenon [@gamsonChangingCulture1987]. Scholars in the field of political communications have proposed that communicators, such as the media or an institution, construct the narrative of a frame for policy positions or public issues in order to activate or restrict a particular response in the intended audience [@panFramingAnalysisApproach1993]. Organized groups of stakeholders can employ similar methods to attract attention to particular issues. Framing can be used to position existing solutions as suitable to address particular issues [@mahFramecriticalPolicyAnalysis2014], which may prevent the public from being aware of other policy approaches that challenge the status quo. The way policy issues are framed is ultimately important to understand because it plays a role in either altering or preserving the existing social perceptions. This, in turn, can affect whether issues are put on the agenda of policy makers and can determine which solutions are proposed to address the problem [@kingdonAgendasAlternativesPublic1984].

Framing of issues is an important step in developing health policy. An obvious example over the past decade is the framing of climate change as a public health issue [e.g., @depouxCommunicatingClimateChange2017; @maibachReframingClimateChange2010; @weathersDevelopmentsFramingClimate2016] to increase public engagement and awareness of the issue. This framing has slowly advanced this issue on public policy agendas as public attention puts pressure on the policy stream to adopt frameworks for action. For example, transport planners also use different frames to guide the extent to which transport policies can be adapted to address climate change. In a recent paper [@reynardGrowthResilienceHow2021], framing analysis was applied to review the representation of issues such as mobility and social exclusion in municipal policies from four western Canadian cities under the current circumstances of climate change. The authors found four primary frames: “The Growing City”, “If You Build It, They Will Come”, “Better City for All”, and a “the Resilient City” [@reynardGrowthResilienceHow2021]. Each frame presented the nature, opportunities, and challenges of climate change in different ways which set the stage for the types of mitigation and adaptation strategies that cities were proposing to address this issue.

In a similar way, we hypothesize that STP stakeholder groups in Ontario have framed AST in particular ways to effectively engage multiple audiences on this policy issue including parents, local residents, school representatives, and municipal representatives. These groups are likely identifying points of intervention and potential opportunities to build support for and encourage AST.

@@

We use topic modelling in R to conduct the framing analysis. Topic modelling is a machine learning technique that can identify what language and concepts are being communicated by analyzing text. This method is more practical for researchers working with large amounts of text because it replaces the manual coding of topics that would normally take place to analyze or summarize textual data [@jacobiQuantitativeAnalysisLarge2016]. In the data pre-processing phase, we tokenize the text in the documents and create a document-term matrix so that it is in the correct format for analysis. We primarily use the following packages: `tidytext` [@R-tidytext], `topicmodels` [@R-topicmodels], `word2vec` [@R-word2vec], and `wordcloud` [@R-wordcloud] to examine text in the documents that were sourced for this project. These packages have functions for determining the frequency of specific words in each document or relationships (e.g., pairs of adjacent terms called bigrams) and correlations between words. These methods can reveal what language and concepts are being communicated to the general public. Topic modelling is a popular method for analyzing text from social media platforms [@albalawiUsingTopicModeling2020] or news articles [@jacobiQuantitativeAnalysisLarge2016]. We estimate latent Dirichlet allocation (LDA) models to classify both the STP and documents according to the topics that are contained within them. This method "treats each document as a mixture of topics, and each topic as a mixture of words" [@silgeTextMining2021]. The model's output is "a set of topics consisting of clusters of words that co-occur in these documents according to certain patterns" [@jacobiQuantitativeAnalysisLarge2016]. Researchers must then interpret the identified topics, as done after other methods of manual coding. We also compare the topics between the policy documents and the academic papers, and use our interpretation to answer the questions outlined above that guide this paper. This methodology is not familiar to many, so we describe in more detail below how we used specific functions and code in R to produce our results.

@@

We analyzed word and document frequency for each set of documents (i.e., corpus). Table \ref{tab:word-table} shows the most frequent terms found in the municipal, transportation consortia, school board, and academic documents. As expected, policy documents and academic papers reference _active travel_, _walking_, _biking_ or _cycling_, and _students_ more than other terms. Each corpus also has _safety_ and _traffic_ as common words which suggests congruence on these key factors between the research literature and how AST is framed to the public. The word _physical_ is present in each corpus, but it's not clear what this refers to (e.g, _physical activity_, _physical health_, or the _physical environment_). Furthermore, documents from STP stakeholder groups discuss _resources_, _information_, and _services_ about school travel. In the section below, the context in which these terms are discussed is explored further.

Unlike the academic papers, policy documents include the words _route_ or _routes_ as frequent terms. This could reflect the role of STP stakeholder groups in identifying safe routes to school to share with parents or families, as well as the STP emphasis on making the physical environment safer for AST. The academic corpus differs from the policy documents in that _parents_ and _distance_ are the second and third most common terms. In addition, _time_, _factors_, _environment_, and _age_ are also identified in the academic papers.  These words are absent from the list of common words in policy documents. Table \ref{tab:word-table} indicates that the academic corpus discusses a broader range of determinants of AST than the policy documents. The number of references for each term in the academic papers is also significantly higher due to the inclusion of more documents.

@@

Bigrams refer to a pair of consecutive words. We use the `unnest_tokens` function from the _tidytext_ package to determine the bigrams. Figures \ref{fig:city-visual}, \ref{fig:consortia-visual}, and \ref{fig:school-visual} show the bigrams that occur more than 5 times for each set of policy documents. These figures help to make further sense of the word frequencies reported above, and highlight the main ideas that are presented to the public in each of the policy corpora. Municipalities primarily discuss _physical activity_ (n = 53) and _public health_ (n = 19) in the context of AST. In addition, _travel planning_ (n = 19), _bike lanes_ (n = 16), and _safe routes_ (n = 14) are also identified, conceivably as either proposed solutions or built environment factors that support AST. Key issues related to transport such as _traffic safety_ (n = 10), _air quality_ (n = 9), and _greenhouse gases_ (n = 9) are conveyed to the public through these policy documents. It is not surprising to find this focus given that municipalities in Ontario are concerned about climate change and have increasingly looked to active modes of travel to offset transport-related emissions in urban areas.

Similar word bigrams are found in school board documents: _travel planning_ (n = 33), _safe routes_ (n = 15), _physical activity_ (n = 10), and _public health_ (n = 10) are among the most common bigrams. Both municipalities and school boards in Ontario seem to emphasize what can be or has been done to improve AST (i.e., policy or planning changes), while outlining some of the benefits of AST at the individual- or community-level to potentially encourage behaviour change (i.e., physical activity for children or improved air quality). Unlike other STP stakeholders, school boards also consider _inclement weather_ (n = 24) and _bus cancellations_ (n = 13). This is likely because many students in Ontario travel to school by bus and this information is presented alongside AST options. Finally, transportation consortia documents highlight topics such as _physical activity_ (n = 10), _pedestrian safety_ (n = 8), _crossing guards_ (n = 6), _travel planning_ (n = 6), and _walk zones_ (n = 6). Biking or cycling is notably absent from transportation consortia documents. Overall, the policy documents There appears convey an emphasis on the built environment, rather than household decision-making.

@@

policy_lda <- LDA(policy_dtm, k = 7, method = "Gibbs", control = NULL, model = NULL)

academic_lda <- LDA(academic_dtm, k = 17, method = "Gibbs", control = NULL, model = NULL)
```
